Efficient ancestry and mutation simulation with msprime 1.0

### 2021-11-19 response-to-reviewers.tex

\begin{reply}
We have removed the some of the examples and shortened the section.
\end{reply}

\begin{point}
As a third example, in the
recombination section after useful comparisons with other approaches, line 354
states that the proposed algorithm is still quadratic in the recombination
rate. The long section from line 354 to 426 details the algorithmic reasons,
and use various analytic approximations and simulations to support this. This
material is of interest only to a few algorithm developers, and I would imagine
it would be better placed in an appendix, with the key observations summarized
in the main text.
\end{point}

\begin{reply}
We have moved this detailed discussion to an appendix.
\end{reply}

\begin{point}
Another feature of the current paper is that a number of topics return several
times with minor variations. This is particularly true for algorithmic
efficiency, and the API, both of which are discussed multiple times. It is of
course true that algorithmic efficiency is the major distinguishing feature of
msprime; nevertheless, the focus of the current paper seems to be the rich
feature set, and the software development model, as well as the various APIs
that allow for a more integrated and less error prone workflow. I would suggest
to introduce separate sections for algorithmic efficiency and the API, and
discuss the various features (as you do now) in their own separate section, but
focus on these features from a user's rather than a developer's perspective.
\end{point}

\begin{reply}
The large number and diversity of features in msprime has made the
organisation of this
manuscript a major challenge, and the present layout of a series of
relatively self-contained topics arrived at after much experimentation.
Issues like interface design and algorithmic efficiency necessarily
cut across most of these topics. Separate sections on efficiency
and API design would need to refer to these other sections for context,
and would lead, we believe, to a longer and less cohesive document.

However, we do agree that some sections are too long and this criticism
has been very helpful to allow us see where content could be cut
without loss of information.
For example, we have cut out references to the API design in the
Simulation Mutations section, and substantially cut back the Demography
section (see next point).
\end{reply}

\begin{point}
Occasionally the paper reads more like documentation than a paper, such as when
the Demography class is introduced and several associated functions are
mentioned - this is not relevant for a paper in my opinion.
\end{point}
\begin{reply}
This is a very helpful observation, and helped us clarify where
text can be cut. We have substantially but back the Demography section.
\end{reply}

@@

\begin{point}
The section about the development model is interesting - very few papers
include such a section, and indeed I think this is one of the stand-out
features of this project that it is so successfully supported by many
contributors. My only suggestion here is to remove the listing of number of
lines of code - this is a very questionable proxy for quality or quantity of
work, and the work is impressive as it is.

\end{point}
\begin{reply}
Thank you for your suggestion, we have removed the number of lines of code
in the revised manuscript.
\end{reply}

\reviewersection

\begin{point}
This manuscript introduces a simulation tool, msprime version 1.0, and
describes features, additions, and performance of the simulator relative to the
earlier msprime software. I will begin by saying that I believe all researchers
in population genomics who use simulations have appreciated the quality work
that has been previously done in the msprime sphere. The contributions proposed
here clearly improve on a number of features, and that of course has value.
\end{point}
\begin{reply}
Thank you, we appreciate the kind words.
\end{reply}

\begin{point}
However, I did not find this work to be appropriate for Genetics, and would
rather view it as more appropriate for a journal like Bioinformatics. The
fundamental reason is that this basically appears to perform old
functionalities faster, and the added functionalities are modest. Again, both
are valuable, but this work doesn't address much larger biological /
statistical inference problems that are in need of development in this area,
which could justify publication in a journal like Genetics.
\end{point}
\begin{reply}
We respectfully disagree, and believe that this paper is both of substantial
interest to the readership of Genetics, and presents a number of significant
innovations that will enable years of research to follow.
These innovations either
allow simulation of models for which there is no other available simulator
(for example, the $\Lambda$-coalescent models and instantaneous bottlenecks)
or at a genuinely transformative performance differential from existing
methods. For example, msprime can simulate E-coli genomes tens of
times faster than specialised bacterial simulators, and selective
sweeps multiple orders of magnitude faster than discoal.
These are not minor incremental improvements, and fundamentally shift
the realms of possibility for population genetic simulations.
\end{reply}

\begin{point}
Specifically, the need of the field is really to simulate more realistic models
in a coalescent framework, rather than the slower forward-in-time choice that
stands as the only option presently. For example, the practical use of a tool
that only simulates strictly neutral histories, or histories with the addition
of positive selection, is likely rather limited empirically. By which I mean,
the functional regions of a genome of course experience purifying selection ,
but there is growing appreciation of the fact that this may result in
background selection effects that may be widespread across the genomes of many
species. In order for the strictly neutral demographic simulations to be of
value, one must identify genomic regions that are not only neutral but are also
unaffected by selection at linked sites. In a great many organisms, such
regions may well not exist at all; in terms of commonly studied species, these
regions may only exist in a handful of large, coding-sparse vertebrate and
plant genomes (though, even if these regions exist, they are not necessarily
easy to identify, or sufficiently large to perform neutral demographic
estimation). I note that the manuscript avoids any mention or discussion of
this problem, and avoids any citations that quantify this issue, but the
neglect doesn't alleviate the problem. The authors should really consider,
discuss, and utilize the following work carefully in this regard:

@@

Castellano et al. 2020. Impact of mutation rate and selection at linked sites
on DNA variation across the genomes of humans and other Homininae. GBE.

Torres et al. 2020. The temporal dynamics of background selection in
nonequilibrium populations. Genetics.

Johri et al. 2021. The impact of purifying and background selection on the
inference of population history: problems and prospects. MBE.

Murphy et al. 2021. Broad-scale variation in human genetic diversity levels is
predicted by purifying selection on coding and non-coding elements. bioRxiv.
\end{point}

\begin{reply}
It is, of course, absolutely correct to point out that neutral simulations
are an approximation, and that this approximation will be highly inaccurate
and inappropriate in many situations. However, it is reasonable to say that
it is both a widely used approximation, and that the
shortcomings of the neutral approximation have been thoroughly discussed
elsewhere. It is surely beyond the scope of this paper
to settle a decades long debate.
\end{reply}

\begin{point}
Relatedly, for the simulations of positive selection, while it has certainly
been common practice in the field for a few decades, it is rather odd to model
a single positively selected site on a background of all other neutral sites.
Positive selection may act on mutations in functional regions, and functional
regions will be largely shaped by purifying selection. Furthermore, the
Hill-Robertson effects between the modeled positively selected mutation and the
neglected array of negatively selected mutations is often non-trivial. In that
sense, the positive selection model is probably only of empirical relevance in
organisms with something approaching free recombination (to eliminate such
linkage effects), as in perhaps HIV or the like. But of course, these sorts of
organisms tend to have small, coding-dense genomes and experience very strong
selective pressures, thus returning to the first problem mentioned above that
regions free of background selection effects may not exist in the first place.
\end{point}
\begin{reply}
This is a fair point and absolutely correct. However, we have simply
implemented the standard approach that has been used and studied extensively
for decades, and have not made any claims beyond this.
\end{reply}

\begin{point}
In summary, I believe this to be a nice software addition that belongs as a
Note in a bioinformatics journal. The software solution that would represent an
important biological advance for Genetics would involve simulating more
realistic models as discussed above (as forward simulators are currently
capable of doing), but at a fast enough rate to be applicable to large-scale
simulation-based genomic inference.
\end{point}

\begin{reply}
Thank you for taking the time to explain your perspective. While we disagree on
the value of this manuscript and its potential interest to the Genetics
readership, we do agree that there is a large gap in present-day simulation
methods and that devising more sophisticated approaches to selection in
particular is a vital avenue for future work.
\end{reply}
\end{document}

### 2021-11-19 paper.tex
Tweak
Update table

Demography
& Improved inferface, import from Newick species tree and
*BEAST~\citep{heled2009bayesian},
Demes support~\citep{gower2022demes}, numerical methods to compute
coalescence rates.\\

@@

\end{tabular}
}
\caption{Major features of \msprime\ 1.0 new since version
0.3.0~\citep{kelleher2016efficient}.}
\end{table}

>>>>

Demography
&
Improved interface with integrated metadata and referencing populations by
name. Import from Newick species tree, *BEAST~\citep{heled2009bayesian},
and Demes~\citep{gower2022demes}. Numerical methods to compute
coalescence rates.\\

@@

\end{tabular}
}
\caption{Major features of \msprime\ 1.0 added since version
0.3.0~\citep{kelleher2016efficient}.}
\end{table}
